Miniproject - Analytic Number Theory

### 2021-12-08 Analytic number theory/submission.tex
Fixed ANT III (iii)

    This is true, as for all $n \in \N$ it holds
        (\chi\log * \chi\mu)(n) &= \sum_{ab = n} \chi(a)\chi(b) \log(a)\mu(b) \\
        &= \sum_{ab = n} \chi(ab) \log(a)\mu(b) = \chi(n) \sum_{ab = n} \log(a) \mu(b) \\
        &= \chi(n) (\log * \mu)(n) = (\chi\Lambda)(n)
Now we want to find an analytic continuation of $L(s, \chi)$ to $\Re(s) > 0$.
First of all, we consider $\chi_0$.
\begin{proposition}[Ex (ii)]
    \label{prop:formula_trivial_L_series}
    For $\Re(s) > 1$ we have
        L(s, \chi_0) = (1 - q^{-s}) \zeta(s)
    In particular, $L(s, \chi_0)$ has a meromorphic continuation to $\Re(s) > 0$ with only one simple pole at $s = 1$.
    As $\chi_0$ is fully multiplicative, we have the Euler product
        L(s, \chi_0) =& \sum_{n \geq 1} \chi_0(n) n^{-s} = \prod_{p \in \primes} \frac 1 {1 - \chi_0(p)p^{-s}} = \prod_{p \neq q} \frac 1 {1 - p^{-s}} \\
        =& (1 - q^{-s}) \prod_{p \in \primes} \frac 1 {1 - p^{-s}} = (1 - q^{-s}) \zeta(s)
    as all products converge absolutely.
For the other Dirichlet characters, the situation is slightly more complicated.
First, we will bound the value of the partial sums of the Dirichlet series for $0 < \Re(s) \leq 1$.
    \label{prop:bound_partsums_character}
    Let $\chi \neq \chi_0$ be a Dirichlet character mod $q$ and consider the sum function
        A(n) := \sum_{1 \leq k \leq n} \chi(k)
    Then $|A(n)| \leq q$ for all $n \in \N$.
    Have
        |A(n)| =& \ \Bigl| \sum_{1 \leq k \leq n} \chi(k) \Bigr| \leq \Bigl| \sum_{q\lfloor n/q \rfloor < l \leq n} \chi(l) \Bigr| + \Bigl| \sum_{0 \leq k < \lfloor n/q \rfloor} \ \underbrace{\sum_{0 < l \leq q} \chi(kq + l)}_{= 0 \ \text{by Prop.~\ref{prop:finite_sums_characters}}} \Bigr| \\
        =& \ \Bigl| \sum_{q\lfloor n/q \rfloor < l \leq n} \chi(l) \Bigr| \leq \sum_{q\lfloor n/q \rfloor < l \leq n} |\chi(l)| = \sum_{q\lfloor n/q \rfloor < l \leq n} 1 \\
        =& q(n/q - \lfloor n/q \rfloor) \leq q
    for all $n \in \N$.
\begin{lemma}[Ex (iii)]
    \label{prop:character_partsums_bounded}
    Let $\chi \neq \chi_0$ be a Dirichlet character.
    Then for $\Re(s) > 0$ have
        \Bigl| \sum_{1 \leq n \leq X} \chi(n) n^{-s} \Bigr| \leq q + \frac {q|s|} {\Re(s)} = O\left( \frac {q|s|} {\Re(s)} \right)
    and in particular, this does not depend on $X$.
    Using partial summation yields
        \sum_{1 \leq n \leq X} \chi(n) n^{-s} =& \sum_{1 - \epsilon < n \leq X} \chi(n) n^{-s} \\
        =& \underbrace{A(1 - \epsilon)}_{= 0} (1 - \epsilon)^{-s} - A(X) X^{-s} + s\int_{1 - \epsilon}^X A(t) t^{-(s + 1)} dt \\
        =& -A(X) X^{-s} + s\int_1^X A(t) t^{-(s + 1)} dt
        \Bigl| \sum_{1 \leq n \leq X} \chi(n) n^{-s} \Bigr| \leq& \ |A(X) X^{-s}| + |s| \int_1^X |A(t) t^{-(s + 1)}| dt \\
        \leq& \ q X^{-\Re(s)} + |s| \int_1^X q t^{-\Re(s) - 1} dt \\
        =& \ q X^{-\Re(s)} + q|s| \left( \frac 1 {\Re(s)} - \frac 1 {X^{\Re(s)}\Re(s)} \right) \\
        \leq& \ q + q \frac {|s|} {\Re(s)} = O\left( \frac {q|s|} {\Re(s)} \right)
    for $\Re(s) > 0$.
Now we can show the analytic continuation of $L(s, \chi)$ to $\Re(s) > 0$.
\begin{proposition}[Ex (iv)]
    Let $\chi \neq \chi_0$ be a Dirichlet character mod $q$.
        L(s, \chi) = s \int_1^\infty A(t) t^{-(s + 1)} dt
    for $\Re(s) > 1$.
    Further, the right-hand side is a holomorphic function on $\Re(s) > 0$ and thus provides an analytic continuation of $L(s, \chi)$ to $\Re(s) > 0$.
    Similar to the proof of Lemma~\ref{prop:character_partsums_bounded}, partial summation yields
        \sum_{n \geq 1} \chi(n) n^{-s} = s \int_1^\infty A(t) t^{-(s + 1)} dt
    By Lemma~\ref{prop:character_partsums_bounded}, this is bounded and hence well-defined and finite for $\Re(s) > 0$.
    Further, the integral converges absolutely by Lemma~\ref{prop:bound_partsums_character}, and thus is holomorphic on $\Re(s) > 0$.
\begin{corollary}[Ex (iv)]
    \label{prop:logarithmic_derivative_L_bounded}
    The function $L(s, \chi)'/L(s, \chi)$ is bounded on a neighborhood of $1$, provided that $L(1, \chi) \neq 0$.
    As $L(s, \chi)'/L(s, \chi)$ is meromorphic on $\Re(s) > 0$, we know that it is holomorphic on some neighborhood of $1$ unless it has a pole at $s = 1$.
    In the third exercise class of ANT, it was shown that this would imply $L(1, \chi) = 0$ or $L(s, \chi)'$ has a pole at $s = 1$.

>>>>

    This is true, as for all $n \in \N$ it holds
    \begin{align*}
        (\chi\log * \chi\mu)(n) &= \sum_{ab = n} \chi(a)\chi(b) \log(a)\mu(b) \\
        &= \sum_{ab = n} \chi(ab) \log(a)\mu(b) = \chi(n) \sum_{ab = n} \log(a) \mu(b) \\
        &= \chi(n) (\log * \mu)(n) = (\chi\Lambda)(n)
    \end{align*}
\end{proof}
Now we want to find an analytic continuation of $L(s, \chi)$ to $\Re(s) > 0$.
First of all, we consider $\chi_0$.
\begin{proposition}[Ex (ii)]
    \label{prop:formula_trivial_L_series}
    For $\Re(s) > 1$ we have
    \begin{equation*}
        L(s, \chi_0) = (1 - q^{-s}) \zeta(s)
    \end{equation*}
    In particular, $L(s, \chi_0)$ has a meromorphic continuation to $\Re(s) > 0$ with only one simple pole at $s = 1$.
\end{proposition}
\begin{proof}
    As $\chi_0$ is fully multiplicative, we have the Euler product
    \begin{align*}
        L(s, \chi_0) =& \sum_{n \geq 1} \chi_0(n) n^{-s} = \prod_{p \in \primes} \frac 1 {1 - \chi_0(p)p^{-s}} = \prod_{p \neq q} \frac 1 {1 - p^{-s}} \\
        =& (1 - q^{-s}) \prod_{p \in \primes} \frac 1 {1 - p^{-s}} = (1 - q^{-s}) \zeta(s)
    \end{align*}
    as all products converge absolutely.
\end{proof}
For the other Dirichlet characters, the situation is slightly more complicated.
First, we will bound the value of the partial sums of a Dirichlet character.
\begin{lemma}[Ex (iii)]
    \label{prop:bound_partsums_character}
    Let $\chi \neq \chi_0$ be a Dirichlet character mod $q$ and consider the sum function
    \begin{equation*}
        A(n) := \sum_{1 \leq k \leq n} \chi(k)
    \end{equation*}
    Then $|A(n)| \leq q$ for all $n \in \N$.
\end{lemma}
\begin{proof}
    Have
    \begin{align*}
        |A(n)| =& \ \Bigl| \sum_{1 \leq k \leq n} \chi(k) \Bigr| \leq \Bigl| \sum_{q\lfloor n/q \rfloor < l \leq n} \chi(l) \Bigr| + \Bigl| \sum_{0 \leq k < \lfloor n/q \rfloor} \ \underbrace{\sum_{0 < l \leq q} \chi(kq + l)}_{= 0 \ \text{by Prop.~\ref{prop:finite_sums_characters}}} \Bigr| \\
        =& \ \Bigl| \sum_{q\lfloor n/q \rfloor < l \leq n} \chi(l) \Bigr| \leq \sum_{q\lfloor n/q \rfloor < l \leq n} |\chi(l)| = \sum_{q\lfloor n/q \rfloor < l \leq n} 1 \\
        =& q(n/q - \lfloor n/q \rfloor) \leq q
    \end{align*}
    for all $n \in \N$.
\end{proof}
Now we can show the analytic continuation of $L(s, \chi)$ to $\Re(s) > 0$.
\begin{proposition}[Ex (iv)]
    Let $\chi \neq \chi_0$ be a Dirichlet character mod $q$.
    Then
    \begin{align*}
        L(s, \chi) = s \int_1^\infty A(t) t^{-(s + 1)} dt
    \end{align*}
    for $\Re(s) > 1$.
    Further, the right-hand side is a holomorphic function on $\Re(s) > 0$ and thus provides an analytic continuation of $L(s, \chi)$ to $\Re(s) > 0$.
\end{proposition}
\begin{proof}
    Using partial summation, we find for some $0 < \epsilon < 1$ that
    \begin{align*}
        \sum_{n \geq 1} \chi(n) n^{-s} =& A(1 - \epsilon) (1 - \epsilon)^{-s} - (-s) \int_{1 - \epsilon}^\infty A(t) t^{-(s + 1)} dt \\
        =& s \int_1^\infty A(t) t^{-(s + 1)} dt
    \end{align*}
    Further, the integral converges absolutely by Lemma~\ref{prop:bound_partsums_character}, and thus is holomorphic on $\Re(s) > 0$.
\end{proof}
\begin{corollary}[Ex (iv)]
    \label{prop:logarithmic_derivative_L_bounded}
    The function $L(s, \chi)'/L(s, \chi)$ is bounded on a neighborhood of $1$, provided that $L(1, \chi) \neq 0$.
\end{corollary}
\begin{proof}
    As $L(s, \chi)'/L(s, \chi)$ is meromorphic on $\Re(s) > 0$, we know that it is holomorphic on some neighborhood of $1$ unless it has a pole at $s = 1$.
    In the third exercise class of ANT, it was shown that this would imply $L(1, \chi) = 0$ or $L(s, \chi)'$ has a pole at $s = 1$.
